---
title: Package Creation and Metadata 🧰
execute: 
  eval: false
---

## Lay the foundations

### Load devtools and create your package infrastructure

```{r}
library(devtools)
create_package("~/Desktop/libminer")
```

### Confirm git is configured correctly (and update if necessary)

This should have been done in your [setup](setup.qmd), but it's not a bad idea to double check.

```{r}
git_sitrep()
```

If the above command returns unsatisfactory information (most importantly, do you have a username and email configured), run the following command to configure git:

```{r}
use_git_config(
  user.name = "Your Name",
  user.email = "Your Email"
)
```

### Initialize this package as a git repo

```{r}
use_git()
```

### Use devtools

We can load devtools automatically via our `.Rprofile` file so that it is always loaded during development

```{r}
use_devtools()
```

Paste the contents of your clipboard in `.Rprofile`, and restart R.

## Write your first function

### Open a new file in `R/` to hold your code

```{r}
use_r("lib-summary")
```

### Write your function definition in this file

```{r}
#| eval: true
lib_summary <- function() {
  pkgs <- utils::installed.packages()
  pkg_tbl <- table(pkgs[, "LibPath"])
  pkg_df <- as.data.frame(pkg_tbl, stringsAsFactors = FALSE)
  names(pkg_df) <- c("Library", "n_packages")
  pkg_df
}
```

### Load your package with `load_all()`

```{r}
load_all()
```

And test out our new function:

```{r}
#| eval: true
lib_summary()
```

## Check your package (and then fix the problems)

```{r}
check()
```

### Set a license

```{r}
use_mit_license()
```

### DESCRIPTION file:

```         
Package: libminer
Title: Explore Your R Libraries
Version: 0.0.0.9000
Authors@R: 
    person("Jane", "Doe",
           email = "jane.doe@something.com", 
           role = c("aut", "cre"),
           comment = c(ORCID = "XXXX-XXXX-XXXX-XXXX"))
Description: Provides functions for learning about your R libraries, and the 
    packages you have installed.
```

### Check again

```{r}
check()
```

### Set a default DESCRIPTION

This avoids having to do this manually every time

```{r}
edit_r_profile()
```

And add this to your `.Rprofile` file

```{r}
# Set usethis options:
options(
  usethis.description = list(
    "Authors@R" = utils::person(
      "Andy", "Teucher",
      email = "andy.teucher@gmail.com",
      role = c("aut", "cre"),
      comment = c(ORCID = "0000-1111-2222-3333")
    )
  )
)

options(
  warnPartialMatchArgs = TRUE,
  warnPartialMatchDollar = TRUE,
  warnPartialMatchAttr = TRUE
)
```

## Use GitHub

```{r}
git_sitrep()

use_github()
```
