---
title: "Documentation 📓"
execute: 
  eval: false
---

### Add a roxygen block just above your function definition

Insert template with `Ctrl + Alt + Shift + R` or `Cmd + Option + Shift + R`

``` r
#' R Library Summary
#'
#' Provides a brief summary of the package libraries on your machine
#'
#' @return A data.frame containing the count of packages in each of the user's
#'   libraries
#' @export
#'
#' @examples
#' lib_summary()
lib_summary <- function() {
  ...
}
```

### Build function documentation (`man/lib_summary.Rd`) from roxygen

```{r}
document()
```

You can preview it with `?function_name`

```{r}
load_all()

?lib_summary
```

### Check package to ensure documentation conforms to standards

```{r}
check()
```

## Create package-level documentation

```{r}
use_package_doc()

document()
```

See rendered documentation at `man/libminer2-package.Rd`

Preview and check again

```{r}
load_all()

?libminer

check()
```

### Install your package

```{r}
install()
library(libminer)

lib_summary() # note one more package than before - that's yours!
```

## README

```{r}
use_readme_rmd()
```

````         
---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# libminer

<!-- badges: start -->

<!-- badges: end -->

The goal of libminer is to provide an overview of your R library setup. It is a toy
package created as a part of a workshop and not meant for serious use.

## Installation

You can install the development version of libminer from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("ateucher/libminer")
```

## Example usage

To get a count of installed packages in each of your library locations, 
optionally with the total sizes, use the `lib_summary()` function:

```{{r example}}
library(libminer)
lib_summary()
# specify `sizes = TRUE` to calculate the total size on disk of your packages
lib_summary(sizes = TRUE)
```
````

```{r}
build_readme()
```

```{r}
check()
```
