---
title: script
format: pdf
execute: 
  eval: false
---

### **Turn on keyboard shortcut broadcasting**

## Package Structure and State

```{r}
.libPaths()
```

***back to slides***

## Get Ready

```{r}
all(c("devtools", "roxygen2", "testthat", "knitr", "pkgdown") %in% installed.packages())
```

```{r}
library(devtools)

has_devel()
```

***back to slides***

## Create a package

```{r}
library(devtools)
create_package("~/Desktop/libminer")
```

*Explore package structure in RStudio with learners*

***back to slides***

```{r}
# gert::git_config()
# gert::git_config_global()

use_git_config(
  user.name = "Andy Teucher", 
  user.email = "andy.teucher@gmail.com"
)
```

```{r}
git_sitrep()
```

Git global: Name, email GitHub user: PAT discovered, User, email(s)

```{r}
use_git()
```

***back to slides***

```{r}
devtools::use_devtools()
```

Restart R - check that devtools is loaded

***back to slides***

```{r}
use_r("lib-summary")
```

```{r}
lib_summary <- function() {
  pkgs <- utils::installed.packages()
  pkg_tbl <- table(pkgs[, "LibPath"])
  pkg_df <- as.data.frame(pkg_tbl, stringsAsFactors = FALSE)
  names(pkg_df) <- c("Library", "n_packages")
  pkg_df
}
```

***back to slides***

```{r}
load_all()

lib_summary()
```

**Commit**

***back to slides***

```{r}
check()
```

***back to slides***

Open DESCRIPTION file

```{r}
use_mit_license()
```

**Commit**

***back to slides***

DESCRIPTION file:

```         
Package: libminer
Title: Explore Your R Libraries
Version: 0.0.0.9000
Authors@R: 
    person("Andy", "Teucher", , "andy.teucher@gmail.com", role = c("aut", "cre"),
           comment = c(ORCID = "0000-0002-7840-692X"))
Description: Provides functions for learning about your R libraries, and the 
    packages you have installed.
```

***back to slides***

```{r}
check()
```

**Commit**

***back to slides***

```{r}
use_github()
```

Go to GitHub page, explore show clone of local

***back to slides***

```{r}
edit_r_profile()
```

*put this in the Discord*

```{r}
# Set usethis options
# 
options(
  usethis.description = list(
    "Authors@R" = utils::person(
      "Jane", "Doe",
      email = "jane@example.com",
      role = c("aut", "cre"),
      comment = c(ORCID = "0000-1111-2222-3333")
    )
  )
)

options(
  usethis.description = list(
    "Authors@R" = utils::person(
      "Andy", "Teucher",
      email = "andy.teucher@gmail.com",
      role = c("aut", "cre"),
      comment = c(ORCID = "0000-0002-7840-692X")
    )
  )
)

options(
  warnPartialMatchArgs = TRUE,
  warnPartialMatchDollar = TRUE,
  warnPartialMatchAttr = TRUE
)
```

***back to slides***

## Documentation

`Ctrl + .`

`Ctrl+Alt+Shift+R`

``` r
#' R Library Summary
#'
#' Provides a brief summary of the package libraries on your machine
#'
#' @return A data.frame containing the count of packages in each of the user's
#'   libraries
#' @export
#'
#' @examples
#' lib_summary()
```

```{r}
document()
```

Go to `man/lib_summary.Rd`

```{r}
load_all()

?lib_summary
```

```{r}
check()
```

Look at NAMESPACE

**commit**

***back to slides***

### Package-level documentation

```{r}
use_package_doc()

document()
```

Go to `man/libminer2-package.Rd`

Preview and check again

```{r}
load_all()

?libminer

check()
```

***back to slides***

```{r}
install()
library(libminer)

lib_summary() # note one more package than before - that's yours!
```

**commit and push**

## Testing

**restart R**

```{r}
use_testthat()
```

Have `R/lib-summary.R` open

```{r}
use_test()
```

```{r}
test_that("lib_summary returns expected results", {
  res <- lib_summary()
  expect_s3_class(res, "data.frame")
  expect_equal(ncol(res), 2)
  expect_equal(names(res), c("Library", "n_packages"))
  expect_type(res$Library, "character")
  expect_type(res$n_packages, "integer")
})

test_that("lib_summary fails appropriately", {
  expect_error(lib_summary("foo"), "unused argument")
})
```

```{r}
test()

check()
```

*reinforce file structure*

**commit**

***back to slides***

## Dependencies

```{r}
use_package("fs")

```

**Look at DESCRIPTION, NAMESPACE (no change)**

**commit**

```{r}
lib_summary <- function(sizes = FALSE) {
  if (!is.logical(sizes)) {
    stop("'sizes' must be logical (TRUE/FALSE).")
  }
  
  pkgs <- utils::installed.packages()
  pkg_tbl <- table(pkgs[, "LibPath"])
  pkg_df <- as.data.frame(pkg_tbl, stringsAsFactors = FALSE)
  names(pkg_df) <- c("Library", "n_packages")

  if (sizes) {
    pkg_df$lib_size <- vapply(
      pkg_df$Library,
      function(x) {
        sum(fs::file_size(fs::dir_ls(x, recurse = TRUE)))
      },
      FUN.VALUE = numeric(1)
    )
  }
  pkg_df
}
```

```{r}
test() # failure for unused argument
```

```{r}
test_that("lib_summary fails appropriately", {
  expect_error(lib_summary(sizes = "foo"), "not interpretable as logical")
})

test_that("sizes argument works", {
  res <- lib_summary(sizes = TRUE)
  expect_equal(names(res), c("Library", "n_packages", "lib_size"))
  expect_type(res$lib_size, "double")
})
```

**commit**

```{r}
check() # will warn about undocumented parameter
```

Ctrl+Alt+Shift+R will insert the spot for the sizes param

```{r}
#' Provides a brief summary of the package libraries on your machine
#'
#' @param sizes Should the sizes of the libraries be calculated?
#'   Logical; default `FALSE`.
#'
#' @return A data.frame containing the count of packages in each of the user's
#'   libraries. A `lib_size` column is included if `sizes = TRUE`.
#' @export
#'
#' @examples
#' lib_summary()
#' lib_summary(sizes = TRUE)
```

```{r}
document()

check()
```

Test it out

```{r}
load_all()

?lib_summary

lib_summary(sizes = TRUE)
```

**commit**

***back to slides***

```{r}
use_import_from("purrr", "map_dbl")
```

*Look at: DESCRIPTION, R/libminer-package.R, NAMESPACE*

```{r}
if (sizes) {
  pkg_df$lib_size <- map_dbl(
    pkg_df$Library,
    ~ sum(fs::file_size(fs::dir_ls(.x, recurse = TRUE))),
  )
}
```

```{r}
test()
```

**Note importance of tests here**

```{r}
check()
```

**commit and push**

## README

```{r}
use_readme_rmd()
```

````         
---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# libminer

<!-- badges: start -->

<!-- badges: end -->

The goal of libminer is to provide an overview of your R library setup. It is a toy
package created as a part of a workshop and not meant for serious use.

## Installation

You can install the development version of libminer from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("ateucher/libminer")
```

## Example usage

To get a count of installed packages in each of your library locations, 
optionally with the total sizes, use the `lib_summary()` function:

```{{r example}}
library(libminer)
lib_summary()
# specify `sizes = TRUE` to calculate the total size on disk of your packages
lib_summary(sizes = TRUE)
```
````

```{r}
build_readme()
```

```{r}
check()

install()
```

**commit**

*return to slides*

## Continuous Integration

```{r}
use_github_action()
```

```{r}
build_readme()
```

**commit and push**

*return to slides*

## pkgdown website

```{r}
use_pkgdown_github_pages()
```

## Vignettes

```{r}
use_vignette("lib-sitrep", "Package Library Situation Report")
```

## Releasing

```{r}
use_release_issue()
```

```{r}
use_news_md()
```

## Bonus - factoring out an internal function

```{r}
#' R Library Summary
#'
#' Provides a brief summary of the package
#' libraries on your machine
#'
#' @param sizes a logical indicating whether or not to calculate
#'    sizes
#'
#' @return A data.frame containing the count
#'    of packages in each of your libraries.
#' @export
#'
#' @examples
#' lib_summary()
#' lib_summary(sizes = TRUE)
lib_summary <- function(sizes = FALSE) {
  pkgs <- utils::installed.packages()
  pkg_tbl <- table(pkgs[, "LibPath"])
  pkg_df <- as.data.frame(pkg_tbl, stringsAsFactors = FALSE)
  names(pkg_df) <- c("library", "n_packages")

  if (sizes) {
    pkg_df <- calculate_sizes(pkg_df)
  }
  pkg_df
}

#' calculate sizes
#'
#' @param df a data.frame
#'
#' @return df with a lib_size column
#' @noRd
calculate_sizes <- function(df) {
  df$lib_size <- map_dbl(
    df$library,
    ~ sum(fs::file_size(fs::dir_ls(.x, recurse = TRUE)))
  )
  df
}
```
